---
title: "HW 10"
author: "Aryss Hearne"
date: "2/11/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
```

```{r}
deserts <- read_csv(here("lab10", "data", "surveys_complete.csv"))
```

```{r}
deserts
```

```{r}
names(deserts)
```

#### 1. Use the function(s) of your choice to get an idea of its structure, including how NA’s are treated. Are the data tidy?

```{r}
# Data look tidy. NAs are treated as NAs. 
```

#### 2. How many genera and species are represented in the data? What are the total number of observations? Which species is most/ least frequently sampled in the study?

```{r}
deserts%>%
  ggplot(aes(x=genus, fill=species))+
  scale_y_log10()+
  geom_bar()+
  coord_flip()+
  labs(title="Observations of Species and Genus Observations",
       x="Genus",
       y="n observations",
       fill="Species Recorded")
```

#### 3. What is the proportion of taxa included in this study? Show a table and plot that reflects this count.

```{r}
deserts%>%
  ggplot(aes(x=taxa))+
  geom_bar()+
  scale_y_log10()
```

#### 4. For the taxa included in the study, use the fill option to show the proportion of individuals sampled by `plot_type`.

```{r}
deserts%>%
  ggplot(aes(x=taxa, fill=plot_type))+
  geom_bar()+
  scale_y_log10()+
  labs(title= "Observations of Taxa in x Plot",
       x="Taxa Recorded",
       y="N Observations",
       fill="Plot Recorded")
```
```{r}
# CUTE!!!!!
```


#### 5. What is the range of weight for each species included in the study? Remove any observations of weight that are NA so they do not show up in the plot.

```{r}
deserts%>%
  ggplot(aes(x=species, y=weight))+
  scale_y_log10()+
  geom_boxplot()+
  coord_flip()
```

#### 6. Add another layer to your answer from #4 using `geom_point` to get an idea of how many measurements were taken for each species.

```{r}
deserts%>%
  ggplot(aes(x=species, y=log10(weight)))+
  geom_point()+
  geom_boxplot()+
  coord_flip()
```

#### 7. Dipodomys merriami is the most frequently sampled animal in the study. How have the number of observations of this species changed over the years included in the study?

```{r}
deserts%>%
  filter(genus=="Dipodomys"& species=="merriami")%>%
  ggplot(aes(x=year))+
  geom_bar()+
  labs(title="Fluctuations of Dipodomys merriami Observations Over Time",
      x="Year",
      Y="N Observations")
```


#### 8. What is the relationship between `weight` and `hindfoot` length? Consider whether or not over plotting is an issue.

```{r}
deserts%>%
  ggplot(aes(x=log10(weight), y=log10(hindfoot_length)))+
  geom_jitter()+
  coord_flip()+
  geom_smooth(method=lm, se=F)+
  labs(title="Weight v Hindfoot_length",
       x="Log10(Weight)",
       y="Log10(Hindfoot Length")
```

```{r}
# That is a VERY ugly graph. ): I would probably restrict it a bit to clean it up - either by taxa or area or... something. It needs an additional constraint. 
```

#### 9. Which two species have, on average, the highest weight? Once you have identified them, make a new column that is a ratio of `weight` to `hindfoot_length`. Make a plot that shows the range of this new ratio and fill by sex.

```{r}
deserts%>%
  group_by(species)%>%
  filter(weight!="NA")%>%
  summarize(mean_weight=mean(weight))%>%
  top_n(2)
  
```

```{r}
deserts%>%
  group_by(species)%>%
  filter(weight!="NA", species=="albigula" | species=="spectabilis")%>%
  mutate(weight_to_hindfoot=(weight/hindfoot_length))%>%
  ggplot(aes(x=species, y=weight_to_hindfoot, shape=sex, color=sex))+
  geom_jitter()+
  coord_flip()+
  labs(title="Average Hindfood Length of Spectabilis and Albigula",
       x="Species",
       y="Ratio of Weight to Hindfoot",
       color="sex")
```

```{r}
# So CUTE!
```


#### 10. Make one plot of your choice! Make sure to include at least two of the aesthetics options you have learned.

```{r}
deserts%>%
  filter(taxa=="Rodent")%>%
  ggplot(aes(x=plot_type, y=weight, shape=sex, color=sex))+
  geom_jitter()+
  coord_flip()+
  labs(title="Where the Rat King At?",
       x="Cribs",
       y="Chonkiness")+
  theme(plot.title = element_text(size = rel(0.95), hjust = 0.5))
  
```

