---
title: "Lab10warmup"
author: "Aryss Hearne"
date: "2/10/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(here)
library(janitor)
```
```{r}
life_history<-readr::read_csv("data/mammal_lifehistories_v2.csv")%>%
  na_if("-999")%>%
  clean_names()
```
```{r}
glimpse(life_history)
```




```{r}
life_history%>%
  count(order,sort=T)
```

```{r}
life_history%>%
  ggplot(aes(x=order))+
  geom_bar()+
  coord_flip()
```

```{r}
life_history%>%
  group_by(order)%>%
  summarize(mean_mass=mean(mass,na.rm=T))
```

```{r}
options(scipen = 999)
```
```{r}
life_history%>%
  group_by(order)%>%
  summarize(mean_mass=mean(mass,na.rm=T))%>%
  ggplot(aes(x=order, y=mean_mass))+
  geom_col()
```

```{r}
life_history%>%
  group_by(order)%>%
  summarize(mean_mass=mean(mass,na.rm=T))%>%
  ggplot(aes(x=order,y=mean_mass))+
  geom_col()+
  coord_flip()+
  scale_y_log10()
```

```{r}
life_history%>%
  ggplot(aes(x=gestation, y=wean_mass))+
  geom_point(na.rm=T)+
  scale_y_log10()+
  geom_smooth(method="lm")
```
```{r}
glimpse(life_history)
```

```{r}
life_history%>%
  ggplot(aes(x=order,y=mass))+
  geom_boxplot()+
  scale_y_log10()+
  coord_flip()
```

```{r}
elephants<-readr::read_csv("data/ElephantsMF.csv")%>%
  clean_names()
```

```{r}
elephants
```

#### 1. On midterm 1, I asked how many male and female elephants were represented in the data. How would you make a plot of this?

```{r}
elephants%>%
  group_by(sex)%>%
  ggplot(aes(x=sex))+
  geom_bar()
```


#### 2. How does the age of elephants compare by sex?

```{r}
elephants%>%
  group_by(sex)%>%
  summarize(mean_age=mean(age, na.rm=T))%>%
  ggplot(aes(x=sex, y=mean_age))+
  geom_col()
```

#### 3. What is the relationship between age and height?

```{r}
elephants%>%
  ggplot(aes(x=age, y=height))+
  geom_point()+
  geom_smooth(method="lm")
```

```{r}
elephants%>%
  ggplot(aes(x=age, y=height))+
  geom_point()+
  geom_smooth(method=lm, se=F)
```

```{r}
elephants%>%
  ggplot(aes(x=age, y=height))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  labs(title="Elephant Age v Height",
       x="Age",
       y="Height (cm)")
```

```{r}
elephants%>%
  ggplot(aes(x=age, y=height))+
  geom_point()+
  geom_smooth(method=lm, se=F)+
  labs(title="Elephants Age v Height",
       x="Age",
       y="Height (cm)")+
  theme(plot.title=element_text(size=12, face="bold"),
        axis.text = element_text(size=10),
        axis.title=element_text(size=10))
```

```{r}
elephants %>% 
  ggplot(aes(x=age, y=height)) + geom_point() + geom_smooth(method=lm, se=F) +
  labs(title = "Elephant Age vs. Height",
       x = "Age",
       y = "Height") +
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

```{r}
elephants %>% 
  ggplot(aes(x=sex, fill=sex))+geom_bar()+
  labs(title = "# Elephants by Sex",
       x = "Sex",
       y = NULL) +
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

```{r}
life_history %>% 
  ggplot(aes(x=gestation, y=log10(mass), size=mass))+
  geom_point(na.rm=T)
```

```{r}
life_history%>%
  filter(gestation>20)
```

```{r}
homerange<-readr::read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")%>%
  clean_names()
```

```{r}
names(homerange)
```
```{r}
homerange
```



#### Make a barplot that shows counts of `ectotherms` and `endotherms`. Label the axes, provide a title, and fill by `thermoregulation`.


```{r}
homerange%>%
  group_by(thermoregulation)%>%
  ggplot(aes(x=thermoregulation, fill=thermoregulation))+
  geom_bar()
```


#### Make a scatterplot that compares log10.mass and homerange. Adjust the size of points by mass. Label the axes and provide a title.

```{r}
homerange%>%
  ggplot(aes(x=mean_hra_m2, y=log10_mass, size=log10_mass))+
  scale_x_log10()+
  geom_point()+ 
  geom_smooth(method="lm")+
  labs(title = "Homerange v Mass",
       x = "Avg Homerange (log10(m2))",
       y = "log10(mass)") +
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
```

#Part 2

```{r}
p<-homerange%>%
  ggplot(aes(x=log10_mass, y=log10_hra))
```

```{r}
p +geom_point(size=1)
```

```{r}
p+geom_point(aes(shape=thermoregulation, color=thermoregulation),
             size=1.75)
```

```{r}
homerange%>%
  count(taxon)
```

```{r}
homerange%>%
  ggplot(aes(x=taxon))+
  geom_bar()+
  coord_flip()+
  labs(title="Observations by Taxon in Homerange Data",
       x="Taxonomic Group")
```

```{r}
homerange%>%
  ggplot(aes(x=taxon, fill=trophic_guild))+
  geom_bar()+
  coord_flip()+
  labs(title="Observations by Taxon in Homerange Data",
       x="Taxonomic Group",
       fill="Trophic Guild")
```

```{r}
homerange%>%
  ggplot(aes(x=taxon, fill=trophic_guild))+
  geom_bar(position="dodge")+
  coord_flip()+
  labs(title="Observations by Taxon in Homerange Data",
       x="Taxonomic Group",
       fill="Trophic Guild")
```

```{r}
homerange%>%
  ggplot(aes(x=taxon, fill=trophic_guild))+
  geom_bar(position="dodge")+
  labs(title="Observations by Taxon in Homerange Data",
       x="Taxonomic Group",
       fill="Trophic Guild")
```

```{r}
homerange%>%
  ggplot(aes(x=taxon, fill=trophic_guild))+
  geom_bar(position=position_fill())+
  scale_y_continuous(labels=scales::percent)+
  coord_flip()
```

```{r}
library(palmerpenguins)
```

```{r}
penguins
```

```{r}
names(penguins)
```

#### 1. Make a bar plot that shows counts of individuals by island. Fill by species, and try both a stacked bar plot and `position="dodge"`.


```{r}
penguins%>%
  group_by(island)%>%
  ggplot(aes(x=island, fill=species))+
  geom_bar()+
  coord_flip()
```

```{r}
penguins%>%
  group_by(island)%>%
  ggplot(aes(x=island, fill=species))+
  geom_bar(position="dodge")+
  coord_flip()
```


#### 2. What are the number of individuals separated by sex on each island?


```{r}
penguins%>%
  group_by(island)%>%
  ggplot(aes(x=island, fill=sex))+
  geom_bar(position="dodge")+
  coord_flip()
```

```{r}
homerange%>%
  ggplot(aes(x=class, y=log10_mass))+
  geom_boxplot()
```

```{r}
homerange%>%
  ggplot(aes(x=class, y=log10_mass, group=taxon))+
  geom_boxplot()
```

```{r}
homerange%>%
  ggplot(aes(x=class, y=log10_mass, fill=taxon))+
  geom_boxplot()
```

