---
title: "HW 11"
author: "Aryss Hearne"
date: "2/15/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
#install.packages("gapminder")
library(gapminder)
library(ggthemes)
library(paletteer)
```

#### 1. Use the function(s) of your choice to get an idea of the overall structure of the data frame, including its dimensions, column names, variable classes, etc. As part of this, determine how NA’s are treated in the data.

```{r}
glimpse(gapminder)
```

```{r}
gapminder%>%
  purrr::map_df(~sum(is.na(.)))
```
```{r}
gapminder_t<-gapminder%>%
  na_if("-999")
```

```{r}
gapminder_t%>%
  purrr::map_df(~sum(is.na(.)))
```
```{r}
#No NA data
```


#### 2. Among the interesting variables in gapminder is life expectancy. How has global life expectancy changed between 1952 and 2007?

```{r}
#gapminder
```

```{r}
#colors<-paletteer::palettes_d_names
#colors
```

```{r}
#install.packages("PNWColors")
#install.packages("devtools")
#devtools::install_github("johannesbjork/LaCroixColoR")
my_lil_palette<- LaCroixColoR::lacroix_palette("Pamplemousse", type = "discrete")
my_lil_palette_cont<-LaCroixColoR::lacroix_palette("Pamplemousse", n = 50, type = "continuous")
barplot(rep(1,14), axes=FALSE, col=my_lil_palette)
```


```{r}
gapminder%>%
  group_by(year)%>%
  summarize(mean_lifeExp=mean(lifeExp))%>%
  ggplot(aes(x=year, y=mean_lifeExp))+
  geom_col(fill="thistle4")+
  theme_solarized()+
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  #scale_fill_manual(values=my_palette_dis)+ Can't get this to work. 
  labs(title="Change in Life Expectancy Over the Years",
       x="Year",
       y="Life Expectancy",
       fill="Life Expectancy")
```


#### 3. How do the distributions of life expectancy compare for the years 1952 and 2007?

```{r}
gapminder%>%
  filter(year=="1952"| year=="2007")%>%
  ggplot(aes(x=country, y=lifeExp, fill=year))+
  geom_col()+
  theme_solarized()+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Countries' Life Expectency in 1952 v 2007",
       x= "Country",
       y= "Life Expectancy",
       fill= "Year")+
  coord_flip()
```
```{r}
gapminder%>%
  filter(year=="1952"| year=="2007")%>%
  ggplot(aes(x=year, y=lifeExp, group=country, color=continent, shape=continent))+
  geom_line(alpha=0.2)+
  theme_solarized()+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Countries' Life Expectency in 1952 v 2007",
       x= "Country",
       y= "Life Expectancy",
       fill= "Year")
```

#### 4. Your answer above doesn’t tell the whole story since life expectancy varies by region. Make a summary that shows the min, mean, and max life expectancy by continent for all years represented in the data.


```{r}
gapminder2<-gapminder%>%
  group_by(continent, year)%>%
  summarize(min_lifeExp=min(lifeExp),
            mean_lifeExp=mean(lifeExp),
            max_lifeExp=max(lifeExp))
gapminder2
```


```{r}
gapminder2%>%
  ggplot(aes(x=year, y=min_lifeExp, group=continent, color=continent))+
  geom_line()+
  geom_point()+
  scale_color_manual(values=my_lil_palette)+
  theme_solarized()+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Minimum Life Expectancy per Continent per Year",
       x="Year",
       y="Life Expectancy",
       fill= "Continent")
  
```

```{r}
gapminder2%>%
  ggplot(aes(x=year, y=max_lifeExp, group=continent, color=continent))+
  geom_line()+
  geom_point()+
  scale_color_manual(values=my_lil_palette)+
  theme_solarized()+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Maximum Life Expectancy per Continent per Year",
       x="Year",
       y="Life Expectancy",
       fill= "Continent")
  
```

```{r}
gapminder2%>%
  ggplot(aes(x=year, y=mean_lifeExp, group=continent, color=continent))+
  geom_line()+
  geom_point()+
  scale_color_manual(values=my_lil_palette)+
  theme_solarized()+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Mean Life Expectancy per Continent per Year",
       x="Year",
       y="Life Expectancy",
       fill= "Continent")
  
```
```{r}
gapminder%>%
  ggplot(aes(x=continent, y=lifeExp, fill=continent))+
  geom_boxplot()+
  scale_fill_manual(values=my_lil_palette)+
  theme_solarized()+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Life Expectancy per Continent from 1952 to 2007",
       x="Continent",
       y="Life Expectancy",
       fill= "Continent")
```

```{r}
#I'm just not sure what kind of data readout this question is looking for. 
```



#### 5. How has life expectancy changed between 1952-2007 for each continent?

```{r}
gapminder%>%
  ggplot(aes(x=year, y=lifeExp, group=continent, color=continent, alpha=0.5))+
  geom_point(alpha=0.5)+
  geom_smooth(se = FALSE, method = "loess")+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Life Expectancy per Continent from 1952 to 2007",
       x="Continent",
       y="Life Expectancy",
       fill= "Continent")
```


#### 6. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer?

```{r}
gapminder%>%
  ggplot(aes(x=lifeExp, y=gdpPercap, color=continent, size=pop))+
  geom_point(alpha=0.5)+
  coord_flip()+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Life Expectancy v GDP",
       x="Life Expectancy",
       y="GDP Per Capita",
       fill= "Continent")
```
#### 7. Which countries have had the largest population growth since 1952?

```{r}
g3<-gapminder%>%
  group_by(country)%>%
  filter(year=="1952" | year=="2007")%>%
  mutate(population_growth=pop-lag(pop,default=first(pop)))
g3
```

```{r}
g3%>%
  arrange(desc(population_growth))
```


#### 8. Use your results from the question above to plot population growth for the top five countries since 1952.



```{r}
gapminder%>%
  filter(country=="China"|country=="India"|country=="United States"|country=="Indonesia"|country=="Brazil")%>%
  ggplot(aes(x=year,y=pop,color=country))+
  geom_point(aes(color=country, size=gdpPercap), alpha=0.5)+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Population over Time by 5 Fastest Growing Countries",
       x="Year",
       y="Population",
       color= "Country")
```
```{r}
gapminder%>%
  filter(country=="China"|country=="India"|country=="United States"|country=="Indonesia"|country=="Brazil")%>%
  ggplot(aes(x=year,y=pop,color=country))+
  geom_point(aes(color=country, size=gdpPercap), alpha=0.5)+
  facet_wrap(~country)+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Population over Time by 5 Fastest Growing Countries",
       x="Year",
       y="Population",
       color= "Country",
       size="GDP per Capita")+
  scale_size(range = c(0.1, 10),
             breaks = 1000000 * c(250, 500, 750, 1000, 1250),
             labels = c("250", "500", "750", "1000", "1250")) 
```


#### 9. How does per-capita GDP growth compare between these same five countries?

```{r}
gapminder%>%
  filter(country=="China"|country=="India"|country=="United States"|country=="Indonesia"|country=="Brazil")%>%
  ggplot(aes(x=year,y=gdpPercap,color=country))+
  geom_point(aes(color=country, size=pop), alpha=0.5)+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="GDP over Time by 5 Countries with the Fastest Growing Populations",
       x="Year",
       y="GDP per Capita",
       color= "Country")+
  scale_size(range = c(0.1, 10),
             guide = "none")
```

```{r}
gapminder%>%
  filter(country=="China"|country=="India"|country=="United States"|country=="Indonesia"|country=="Brazil")%>%
  ggplot(aes(x=year, y=gdpPercap, color=country))+
  geom_point(aes(color=country, size=pop), alpha=0.5)+
  facet_wrap(~country)+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="GDP over Time by 5 Countries with the Fastest Growing Populations",
       x="Year",
       y="GDP per Capita",
       color= "Country",
       size="Population")+
  scale_size(range = c(0.1, 10),
             guide = "none")

  
```


#### 10. Make one plot of your choice that uses faceting!

```{r}
gapminder%>%
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop))+
  geom_point(alpha=0.5)+
  geom_text(aes(x=gdpPercap, y=lifeExp+3, label=country),
            color="grey50",
            data=filter(gapminder, pop>1000000000  | country %in% c("Nigeria", "United States")))+
  scale_x_log10(limits=c(200,60000))+
  scale_size(range=c(0.1,10),
             guide="none")+
  facet_wrap(~year)+
  scale_color_manual(values=my_lil_palette)+ 
  theme_solarized()+
  theme(legend.position="top",
        axis.line = element_line(color = "grey85"),
        axis.ticks = element_line(color = "grey85"),
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Global Life Expectancy v GDP (1952-2007)",
       x="GDP per Capita",
       y="Life Expectancy",
       color= "Continent")
```



