---
title: "Lab11 warmups"
author: "Aryss Hearne"
date: "2/15/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#intall.packages("RColorBrewer")
#install.packages("paletteer")
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
options(scipen=999)
```

```{r}
deserts<-readr::read_csv("data/surveys_complete.csv")
homerange<-readr::read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```
```{r}
deserts%>%
  filter(species_id=="DM"|species_id=="DS")%>%
  group_by(year,species_id)%>%
  summarise(n=n())
```
```{r}
deserts$year<-as.factor(deserts$year)
```
```{r}
deserts%>%
  filter(species_id=="DM"|species_id=="DS")%>%
  group_by(year, species_id)%>%
  summarize(n=n())%>%
  ggplot(aes(x=year,y=n,group=species_id,color=species_id))+
  geom_line()+
  geom_point(shape=1)+
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  labs(title="Number of Samples for Species DM",
       x="Year",
       fill="n")
  
```

```{r}
homerange%>%
  ggplot(aes(x=log10.mass))+
  geom_histogram(alpha=0.6,color="black",fill="deeppink4", bins=40)+
  labs(title="Distribution of Body Mass")
```

```{r}
grDevices::colors()
```


```{r}
homerange%>%
  ggplot(aes(x=log10.mass))+
  geom_density(fill="thistle3", alpha=0.6, color="black")+
  labs(title="distribution of Body Maxx")
```

```{r}
homerange%>%
  ggplot(aes(x=log10.mass))+
  geom_histogram(aes(y=..density..),binwidth=.4,fill="thistle3",alpha=0.8,color="deeppink4")+
  geom_density(color="slateblue")+
  labs(title="Distribution of Body Maxx")
```

#Practice

#### 1.Make a histogram of `log10.hra`. Make sure to add a title.

```{r}
homerange%>%
  ggplot(aes(x=log10.hra))+
  geom_histogram(aes(y=..density..), binwidth =.4, fill="seagreen2", alpha=0.8, color="deeppink4")+
  labs(title="Distribution of Homerange")
```

#### 2. Now plot the same variable using `geom_density()`.

```{r}
homerange%>%
  ggplot(aes(x=log10.hra))+
  geom_density(fill="seagreen2", alpha=0.5, color="olivedrab")+
  labs(title="Distribution of Homerange")
```

#### 3. Combine them both!

```{r}
homerange%>%
  ggplot(aes(x=log10.hra))+
  geom_histogram(aes(y=..density..),binwidth = 0.4, fill="seagreen2",alpha=0.6, color="olivedrab")+
  geom_density(fill="palegreen4", alpha=0.45, color="wheat4")
```

```{r}
glimpse(homerange)
```

```{r}
homerange
```

```{r}
homerange%>%
  select(log10.mass)%>%
  summarize(min=min(log10.mass),
            max=max(log10.mass))
```

```{r}
homerange<-homerange%>%
  mutate(mass_category=case_when(log10.mass<=1.75~"small",
         log10.mass>1.74&log10.mass<=2.75~"medium",
         log10.mass>2.75~"large"))
```

```{r}
homerange%>%
  ggplot(aes(x=mass_category,fill=trophic.guild))+
  geom_bar()+
  coord_flip()+
  labs(title="observations by Taxon and Mass Category in Homerange Data",
       x="Mass Category",
       fill="Trophic Guild")
```

# Practice

#### Use `case_when()` to make a new column `range_category` that breaks down `log10.hra` into very small, small, medium, and large classes based on quartile. 


```{r}
names(homerange)
```


```{r}
homerange%>%
  select(log10.hra)%>%
  summarize(min=min(log10.hra),
            max=max(log10.hra))
```

```{r}
homerange<-homerange%>%
  mutate(range_category=case_when(log10.hra<=2.768~"very small",
                                  log10.hra>2.768&log10.hra<=5.5366045~"small",
                                  log10.hra>5.5366045&log10.hra<=8.305~"medium",
                                  log10.hra>8.305&log10.hra<=9.55033~"large"))
```


```{r}
#homerange
```

#### 2. How many and which taxonomic classes are represented in each of the `range_category` classes?

```{r}
homerange%>%
  ggplot(aes(x=range_category, fill=taxon))+
  geom_bar()+
  coord_flip()+
  labs(title="Observations by Txon in Range Category in Homerange Data",
       x="Range Category",
       fill="Taxon")
```

#### 3. Isolate the small `range_category` and plot the range of `log10.mass` by taxonomic class.

```{r}
homerange%>%
  filter(range_category=="small")%>%
  ggplot(aes(x=log10.mass, y=taxon))+
  geom_col()+
  coord_flip()+
  labs(title="Taxa with Small Range Sorted by Mass",
       y="Taxon",
       x="Mass")
```

#Part 2

```{r}
p<-homerange%>%
  ggplot(aes(x=taxon, fill=trophic.guild))+
  geom_bar(na.rm=T,position="dodge")
p
```

```{r}
p+theme_linedraw()+
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  labs(title="Observations by Taxon in Homerange Data",
       x= NULL,
       y="n",
       fill="Trophic Guild")
```

```{r}
p+theme_linedraw()+
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  labs(title="Observations by Taxon in Homerange Data",
       x=NULL,
       y="n",
       fill="Trophic Guild")
```

```{r}
#install.packages("ggthemes")
library(ggthemes)
```


```{r}
ls("package:ggthemes")[grepl("theme_", ls("package:ggthemes"))]
```

```{r}
p+
  theme_pander()+
  theme(legend.position = "bottom",
        axis.text.x=element_text(angle=60,hjust=1))+
  labs(title="Observations by Taxon in Homerange Data",
       x= NULL,
       y="n",
       fill="Trophic Guild")
```

```{r}
?RColorBrewer
```

```{r}
display.brewer.pal(8,"PRGn")
```

```{r}
p+
  scale_fill_brewer(palette="Paired")+
  theme(legend.position="bottom",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Observations by Taxon in Homerange Data",
       x=NULL,
       y="n",
       fill="Trophic Guild")
```

```{r}
colors<-paletteer::palettes_d_names
```

```{r}
my_palette <- paletteer_d("ggsci::hallmarks_light_cosmic")
```

```{r}
barplot(rep(1,14), axes=FALSE, col=my_palette)
```

```{r}
p+scale_fill_manual(values=my_palette)+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 60, hjust=1))+
  labs(title = "Observations by Taxon in Homerange Data",
       x = NULL,
       y= "n",
       fill= "Trophic Guild")
```


```{r}
homerange%>%
  ggplot(aes(x=log10.mass, y=log10.hra, color=locomotion))+
  geom_point()+
  xlim(0,4)+
  ylim(1,6)
```

