---
title: "Lab 8 HW"
author: "Aryss Hearne"
date: "2/3/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
```
```{r}
#install.packages("here")
```
```{r}
library(here)
```
```{r}
sydneybeaches<-read_csv(here("lab8", "data", "sydneybeaches.csv"))%>%
  janitor::clean_names()
```

#### 2. Are these data “tidy” per the definitions of the tidyverse? How do you know? Are they in wide or long format?

```{r}
sydneybeaches
```

```{r}
#This is in a long format and it looks like the data are currently tidy (except maybe the date as it's a concatonated form of month-day-year. 
``` 


#### 3. We are only interested in the variables site, date, and enterococci_cfu_100ml. Make a new object focused on these variables only. Name the object `sydneybeaches_long`

```{r}
sydneybeaches_long<-
  sydneybeaches%>%
  select(site, date, enterococci_cfu_100ml)
sydneybeaches_long
```

#### 4. Pivot the data such that the dates are column names and each beach only appears once. Name the object sydneybeaches_wide

```{r}
sydneybeaches_wide<-
sydneybeaches_long%>%
  pivot_wider(names_from = "date",
              values_from="enterococci_cfu_100ml")
sydneybeaches_wide  
```

#### 5. Pivot the data back so that the dates are data and not column names.

```{r}
sydneybeaches_long2<-
  sydneybeaches_wide%>%
  pivot_longer(-site,
               names_to="date",
               values_to="enterococci_cfu_100ml")
sydneybeaches_long2
```

#### 6.We haven’t dealt much with dates yet, but separate the date into columns day, month, and year. Do this on the sydneybeaches_long data.

```{r}
sydneybeachdate<-
sydneybeaches_long%>%
  separate(date, into=c("month","day","year"),
           sep="/")
sydneybeachdate
```

#### 7. What is the average `enterococci_cfu_100ml` by year for each beach. Think about which data you will use- long or wide.

```{r}
sydneybeachdate%>%
  group_by(site,year)%>%
  summarize(mean_enterococci=mean(enterococci_cfu_100ml, na.rm=T))%>%
  arrange(year)
```


#### Make the output from question 7 easier to read by pivoting it to wide format.

```{r}
sydneybeachdate_avg<-
  sydneybeachdate%>%
  group_by(site,year)%>%
  summarize(mean_enterococci=mean(enterococci_cfu_100ml, na.rm=T))%>%
  arrange(year)
sydneybeachdate_avg
```

```{r}
sydneybeachdate_avg_wide<-
  sydneybeachdate_avg%>%
  pivot_wider(names_from = "year",
              values_from = "mean_enterococci")
sydneybeachdate_avg_wide
```

#### 9. What was the most polluted beach in 2018?
 
```{r}
sydneybeachdate_avg_wide%>%
  select(site, "2018")%>%
  arrange(desc(`2018`))
```
 
 