---
title: "Bis 15L | Lab 1 Homework"
author: "Aryss Hearne | 914641636"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### 1. Please complete the **[class survey](https://forms.gle/AHHXd3aobaAdkkFg9)** if you have not already done so.

*Complete!*

##### 2. What is the difference between R and RStudio? What is GitHub and why is it useful to programmers?

*R is merely a programming language. It can do stuff (like arithmetic) - but not a ton of stuff. At least, not on its own. This is where Rstudio comes in. Rstudio is a GUI which allows users to more easily manipulate R for their intended purposes.* 

*GitHub is kind of like an instagram for coders. It helps to connect folk and allow them to more easily share/ collaborate on code-related projects.* 

##### 3. Navigate to my **[BIS15L](https://github.com/jmledford3115/BIS15LW2021_jledford)** on GitHub.Notice that near the bottom there is a nice, clean description of the repository. You want this on your repository! It is built by making edits to the `README.md` file using the RMarkdown conventions. Edit your repository’s `README.md` file so that you have a clean informative description, including your email address. Be creative! You don’t need to just copy mine. There are lots of examples online. 

*Done!*

##### 4. Calculate the following expressions. Be sure to include each one in a separate code chunk.
```{r}
+5-3*2
```
```{r}
+8/2**2
```
##### 5. Did any of the results in #4 surprise you? Write two programs that calculate each expression such that the result for the first example is 4 and the second example is 16.

*No, I don't think so. Order of operations (PEMDAS) gets us to -1 in the first eqn. And the second makes sense since 2**2 is (what I imagine) shorthand for a power operation.*

##### 6. `Objects` in R are a way in which we can store data or operations. We will talk more about objects next week. For now, make a new object `pi` as 3.14159265359 by running the following code chunk. You should now see the object `pi` in the environment window in the top right.

```{r}
pi <- 3.14159265359
```

*Cool! Done.*

##### 7. Let’s say we want to multiply `pi` by 2. Using the same arithmetic principles that we just learned, write a code chunk that performs this operation using the object we created.

```{r}
(+5-3*2)*pi
```
```{r}
(+8/2**2)*pi
```

##### 8. In order to get help with any command in R, just type a `?` in front the command of interest. Practice this by running the following code chunk.

```{r}
?mean
```

*it came up with this*

> ### **Arithmetic Mean**
>
> #### **Description**
> 
>Generic function for the (trimmed) arithmetic mean.
>
> #### **Usage**
>
> mean(x, ...)
>
> ##Default S3 method:
> 
> mean(x, trim = 0, na.rm = FALSE, ...)
>
> ...

##### 9. Let’s calculate the mean for the numbers 2, 8, 6, 4, 9, 10. I have built an object `x` for you below so all you need to do is run the first code chunk and then create a second code chunk that shows the calculation. Give it a try!

```{r}
x <- c(2, 8, 6, 6, 7, 4, 9, 9, 9, 10)

mean(x)
```
##### 10. Repeat the procedure above, but this time calculate the median.

```{r}
median(x)
```



