---
title: "Lab12_HW"
author: "Aryss Hearne"
date: "2/24/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(here)
library(ggmap)
library(albersusa)
library(ggthemes)
```


#### 1. The first data set represent sightings of grizzly bears (Ursos arctos) in Alaska.
#### 2. The second data set is from Brandell, Ellen E (2021), Serological dataset and R code for: Patterns and processes of pathogen exposure in gray wolves across North America, Dryad, Dataset.

```{r}
see_bears<-readr::read_csv("data/bear-sightings.csv")%>%tidy_names()
see_wolves<-readr::read_csv("data/wolves_data/wolves_dataset.csv") #Not easily tidyable
```

#### 3. Load the `grizzly` data and evaluate its structure. As part of this step, produce a summary that provides the range of latitude and longitude so you can build an appropriate bounding box.

```{r}
names(see_bears)
```
```{r}
summary(see_bears)
```
```{r}
anyNA(see_bears)
```
```{r}
see_bears%>%
  select(latitude, longitude)%>%
  summary()
```

#### 4. Use the range of the latitude and longitude to build an appropriate bounding box for your map.
#### 5. Load a map from stamen in a terrain style projection and display the map.

```{r}
latb<-c(55.02, 70.37)
longb<-c(-166.2, -131.3)
bboxb<-make_bbox(longb,latb, f=0.05)
mapb<-get_map(bboxb, maptype = "terrain", source="stamen")
ggmap(mapb)
```

#### 6. Build a final map that overlays the recorded observations of grizzly bears in Alaska. 

```{r}
ggmap(mapb)+
  geom_point(data=see_bears, aes(x=longitude, y=latitude))+
  labs(x = "Longitude", y = "Latitude", title = "Grizzly Sightings")+
  theme_solarized()+
  theme(legend.position="right",
        axis.text.x=element_text(angle=60, hjust=1))
```

#### 7. Let’s switch to the wolves data. Load the data and evaluate its structure.

```{r}
#names(see_wolves) #needs to be cleaned
#summary(see_wolves)
#anyNA(see_wolves) #Yes
see_wolves
```

#### 8. How many distinct wolf populations are included in this study? Mae a new object that restricts the data to the wolf populations in the lower 48 US states.

```{r}
n_distinct(see_wolves$pack.size)
```
```{r}
see_wolves_below<-see_wolves%>%
  filter(lat<=48)%>%
  group_by(pack.size)
see_wolves_below
```

#### 9. Use the `albersusa` package to make a base map of the lower 48 US states.

```{r}
library(albersusa)
us_comp <- usa_sf()
US_comp_lower<-us_comp%>%
  filter(name!="Alaska", name!="Hawaii")
```

#### 10. Use the relimited data to plot the distribution of wolf populations in the lower 48 US states.


```{r}
ggplot()+
  geom_sf(data=US_comp_lower,size=0.125, alpha=0.25)+
  geom_point(data=see_wolves_below, aes(long, lat, color=pop, size=standard.packsize), alpha=0.5)+
  theme_solarized_2()+
  theme(legend.position="right",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Wolf Pack Sighting",
       x="Longitude",
       y="Latitude",
       size="Standard Packsize",
       color="Population")
  
```

#### 11. What is the average pack size for the wolves in this study by region?

```{r}
see_wolves_below%>%
  group_by(lat)%>%
  summarize(mean_packsize=mean(pack.size, na.rm=T))
```

#### 12. Make a new map that shows the distribution of wolves in the lower 48 US states but which has the size of location markers adjusted by pack size.

```{r}
ggplot()+
  geom_sf(data=US_comp_lower,size=0.125, alpha=0.25)+
  geom_point(data=see_wolves_below, aes(long, lat, color=pop, size=pack.size), alpha=0.5)+
  theme_solarized_2()+
  theme(legend.position="right",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Wolf Pack Sighting",
       x=NULL,
       y=NULL,
       size="Packsize",
       color="Population")
```


