---
title: "Lab 12 Warmups"
author: "Aryss Hearne"
date: "2/23/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(ggmap)
```

```{r}
files<-list.files("data/spiders",".csv")
files
```
```{r}
files<-list.files("data/spiders",".csv",full.names=TRUE)
files
```

```{r}
spider_list<-lapply(files, read_csv)
spider_list[[3]]
```
```{r}
class(spider_list[[3]])
```
#### 1. What are the names of our list elements?

```{r}
names(spider_list[[3]])
```
```{r}
names<-list.files("data/spiders",".csv")
names_list<-strsplit(names,split=".csv")
names_list
```
```{r}
names_vec<-unlist(names_list)
names_vec
```

```{r}
names(spider_list)<-names_vec
names(spider_list)
```
#### 1. Now that our list elements are named, how could we access the Butte County data by name?


```{r}
spider_list$Butte
```

```{r}
spiders_all<-bind_rows(spider_list)
spiders_all
```

```{r}
table_A<-readr::read_csv("data/table_A.csv")
table_B<-readr::read_csv("data/table_B.csv")
```
```{r}
inner_ex<-inner_join(table_A,table_B,by="customer_ID")
inner_ex
```

```{r}
left_ex<-left_join(table_A,table_B,by="customer_ID")
left_ex
```

```{r}
right_ex<-right_join(table_A,table_B,by="customer_ID")
right_ex
```
```{r}
full_ex<-full_join(table_A,table_B,by="customer_ID")
full_ex
```
```{r}
semi_ex<-semi_join(table_A,table_B,by="customer_ID")
semi_ex
```
```{r}
anti_ex<-anti_join(table_A,table_B,by="customer_ID")
anti_ex
```

```{r}
spiders_loc <- read_csv("data/spiders locations/spiders_locations.csv")
spiders_loc
```
```{r}
spiders_with_locs<-left_join(spiders_all, spiders_loc,by=c("Accession"))
summary(spiders_with_locs)
```


```{r}
class(spiders_with_locs$Date)
```

```{r}
day<-today()
str(day)
```

```{r}
datetime<-now()
datetime
```
```{r}
dmy(spiders_with_locs$Date)
```

#### 1. Convert each of the examples above using lubridate.

```{r}
year(day)
```
```{r}
year(datetime)
```

```{r}
month(day)
```
```{r}
day(day)
```
```{r}
hour(datetime)
```
```{r}
minute(datetime)
```
```{r}
second(datetime)
```
```{r}
wday(datetime, label=T)
```

```{r}
write.csv(spiders_with_locs, file="spiders_with_locs.csv", row.names=F)
```

```{r}
#install.packages("remotes")
#remotes::install_github("https://github.com/hrbrmstr/albersusa")
```


# Part 2


```{r}
library(ggmap)
```

```{r}
spiders<-readr::read_csv("data/spiders_with_locs.csv")%>%
  clean_names()
```

```{r}
spiders<-spiders%>%filter(latitude<=42)
```

```{r}
spiders%>%
  select(latitude, longitude)%>%
  summary()
```

```{r}
lat<-c(34.67,41.80)
long<-c(-124.1, -115.5)
bbox<-make_bbox(long,lat,f=0.05)
```

```{r}
map1<-get_map(bbox, maptype="terrain",source="stamen")
```
```{r}
ggmap(map1)
```

```{r}
ggmap(map1)+
  geom_point(data=spiders,aes(longitude,latitude))+
  labs(x="Longitude",
       y="Latitude",
       title="Spider Locations")
```


#### 1. Map the spider locations with a different type of base map. Try to adjust the map using different aesthetics including `size`, `color`, and `alpha`.

```{r}
names(spiders)
```


```{r}
map2<-get_map(bbox, maptype="watercolor",source="stamen")
library(ggthemes)
my_lil_palette<- LaCroixColoR::lacroix_palette("Pamplemousse", type = "discrete")
my_lil_palette_cont<-LaCroixColoR::lacroix_palette("Pamplemousse", n = 50, type = "continuous")
barplot(rep(1,14), axes=FALSE, col=my_lil_palette)
```

```{r}
ggmap(map2)+
  geom_point(data=spiders,aes(longitude,latitude, color=county, size=accession), alpha=0.5)+
  labs(x="Longitude",
       y="Latitude",
       title="Spider Locations")+
  scale_color_manual(values=my_lil_palette_cont)+
  theme_solarized()+
  theme(legend.position="right",
        axis.text.x=element_text(angle=60, hjust=1))+
  labs(title="Spider Locations",
       x=NULL,
       y=NULL,
       color="County",
       size="Accension")+
  scale_size(range = c(1, 10))

```

```{r}
library(albersusa)
```

```{r}
us_comp<-usa_sf()
```
```{r}
cnty_comp<-counties_sf()
```
```{r}
ggplot()+
  geom_sf(data=us_comp, size=0.125)+
  theme_linedraw()+
  labs(title="USA State Boundaries")
```
```{r}
ca_comp<-us_comp%>%
  filter(name=="California")
```

```{r}
ca_cnty_comp<-cnty_comp%>%
  filter(state=="California")
```

```{r}
ggplot()+
  geom_sf(data=ca_comp, size=0.125)+
  geom_sf(data=ca_cnty_comp, size=0.125)
```

```{r}
ggplot()+
  geom_sf(data=ca_comp, size=0.125)+
  geom_sf(data=ca_cnty_comp, size=0.125)+
  geom_point(data=spiders, aes(longitude, latitude))
```

```{r}
ggplot() +
  geom_sf(data = ca_comp, size = 0.125)+
  geom_sf(data = ca_cnty_comp, size = 0.125)+
  geom_point(data = spiders, aes(longitude, latitude))+
  labs(x = "Figure 1: Distribution of Usofila spiders")+
  theme_minimal() +
  theme(axis.text = element_blank()) +
  theme(axis.title.x = element_text(hjust=0.5, size = 10)) +
  theme(axis.title.y = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = c(0.9, 0.3)) -> uso_map
uso_map
```


#Practice
#### Make a new column that labels each specimen as occurring in a cave or on the surface. The word “Cave” is part of the locality information.

```{r}
spiders
```

```{r}
spiders<-spiders%>%
  mutate(depth=case_when(grepl("Cave", locality)~ "Subterranian",
                         grepl(".", locality)~ "Terranian"))
```

```{r}
ggplot() +
  geom_sf(data = ca_comp, size = 0.125)+
  geom_sf(data = ca_cnty_comp, size = 0.125)+
  geom_point(data = spiders, aes(longitude, latitude, color=depth))+
  labs(x = "Figure 1: Distribution of Usofila spiders")+
  theme_minimal() +
  theme(axis.text = element_blank()) +
  theme(axis.title.x = element_text(hjust=0.5, size = 10)) +
  theme(axis.title.y = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = c(0.9, 0.3))
```

