---
title: "Homework 2"
author: "Aryss Hearne | 915651636"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### 1. What is a vector in R? 

*A vector is an organizational tool for different types of data within R. These data include characters, integers, and numeric data.*

##### 2. WHat is a data matrix in R?

*Data matrices are ways of concatenating large pieces of data. They stack vectors similar to data tables.*

##### 3. Below are data collected by three scientists (Jill, Steve, Susan in order) measuring temperatures of eight hot springs. Run this code chunk to create the vectors.

```{r}
spring_1 <- c(36.25, 35.40, 35.30)
spring_2 <- c(35.15, 35.35, 33.35)
spring_3 <- c(30.70, 29.65, 29.20)
spring_4 <- c(39.70, 40.05, 38.65)
spring_5 <- c(31.85, 31.40, 29.30)
spring_6 <- c(30.20, 30.65, 29.75)
spring_7 <- c(32.90, 32.50, 32.80)
spring_8 <- c(36.80, 36.45, 33.15)
```

##### 4. Build a data matrix that has the springs as rows and the columns as scientists.

```{r}
Springn <- c(spring_1,spring_2,spring_3, spring_4, spring_5, spring_6, spring_7, spring_8)
spring_matrix <- matrix(Springn, nrow=8, byrow=T)
spring_matrix
```

##### 5. The names of the springs are 1.Bluebell Spring, 2.Opal Spring, 3.Riverside Spring, 4.Too Hot Spring, 5.Mystery Spring, 6.Emerald Spring, 7.Black Spring, 8.Pearl Spring. Name the rows and columns in the data matrix. Start by making two new vectors with the names, then use `colnames()` and `rownames()` to name the columns and rows.

```{r}
scientist <- c("Jill", "Steve", "Susan")
Springs <- c("Bluebell Spring","Opal Spring","Riverside Spring","Too Hot Spring", "Mystery Spring", "Emerald Spring", "Black Spring", "Pearl Spring")
colnames(spring_matrix) <- scientist
rownames(spring_matrix) <- Springs
spring_matrix
```

##### 6. Calculate the mean temperature of all eight springs.

```{r}
Bluebell <- spring_matrix[1,]
Bluebell_Mean <- mean(Bluebell)
Opal <- spring_matrix[2,]
Opal_mean <- mean(Opal)
Riverside <- spring_matrix[3,]
Riverside_mean <- mean(Opal)
Hot <- spring_matrix[4,]
Hot_mean <- mean(Opal)
Mystery <- spring_matrix[5,]
Mystery_mean <- mean(Opal)
Emerald <- spring_matrix[6,]
Emerald_mean <- mean(Opal)
Black <- spring_matrix[7,]
Black_mean <- mean(Opal)
Pearl <- spring_matrix[8,]
Pearl_mean <- mean(Opal)
Mean_Temp <- c(Bluebell_Mean,Opal_mean,Riverside_mean,Hot_mean,Mystery_mean,Emerald_mean, Black_mean, Pearl_mean)
Mean_Temp
```
##### 7. Add this as a new column in the data matrix.

```{r}
spring_matrix <- cbind(spring_matrix,Mean_Temp)
spring_matrix
```
##### 8. Show Susan’s value for Opal Spring only.

```{r}
Susan_Opal <- spring_matrix[2,3]
Na <- NA
SusanOpal <- c(Na,Na,Susan_Opal,Na)
spring_matrix <- rbind(spring_matrix,SusanOpal)
spring_matrix
```
##### 9. Calcualte the mean for Jill's column only

```{r}
Jillstemp <- spring_matrix[,1]
Jill_Mean_temp <- mean(Jillstemp, na.rm=T)
Mean_Temp_Jill <- c(Jill_Mean_temp, Na, Na, Na)
spring_matrix <- rbind(spring_matrix,Mean_Temp_Jill)
spring_matrix
```

##### 10. Use the data matrix to perform one calculation or operation of your interest.

```{r}
G_T <- mean(Mean_Temp, na.rm=T)
Avg_temp <- c(Na, Na, Na, G_T)
spring_matrix <- rbind(spring_matrix, Avg_temp)
spring_matrix
```

