---
title: "Lab2_1 AH"
author: "Aryss Hearne"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
x <- 42
```
```{r}
x
```
```{r}
treatment <- 36
control <- 38
my_experiment <- sum(treatment,control)
my_experiment
```
```{r}
e <- 12
```
```{r}
my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```

```{r}
class(my_numeric)
```
```{r}
is.integer(my_numeric) #is My-numeric an integer?
```
```{r}
my_integer <- 
  as.integer(my_numeric) #create a new object specified as an integer
is.integer(my_integer) #is my_numeric an integer?
```

```{r}
my_missing <- NA
is.na(my_missing)
```
```{r}
anyNA(my_missing)
```
#Practice

```{r}
new_vector <- c(7, 6.2, 5, 9, NA, 4, 9.8, 7, 3, 2)
mean(new_vector)
```
##### 3. How do you interpret this result? What does this mean about NAs?

*Seems like NAs will void an entire set of numbers if you're trying to arithmetically manipulate the set.*

##### 4. Recalculate the mean using the following code chunk. Why is the useful?

```{r}
mean(new_vector, na.rm=T) #na.rm removes the NA values in the vector
```

*this is super helpful because if you're working with a set of numbers you know to have a NA or, alternatively, if you're working with a set of numbers that's large enough that there's ambigiouty around whether or not there's an NA, you can simply run this code to nixx any NAs which might cause issues.* 


