---
title: "Lab 2_2 AH"
author: "Aryss Hearne | 914641636"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vectors
```{r}
my_vector <- c(10,20,30)
days_of_the_week <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
my_vector_sequence <- c(1:100)
my_vector_sequence
```

```{r}
days_of_the_week[3]
```
## Practice

##### 1. Evaluate all numerics `in my_vector_sequence` to determine which has a value of 15.

```{r}
my_vector_sequence[my_vector_sequence=15]
```
##### 2. We can use operators such as <, >, ==, <==, etc. Show all values in `my_vector_sequence` that are less than 10. How about less than or equal to 10?

```{r}
my_vector_sequence[my_vector_sequence <=9]

my_vector_sequence[my_vector_sequence <=10]
```
##### 3. If you use `[]` then you only get the values, not the logical evaluation of the entire vector. Experiment with this by adjusting the chunk below.
```{r}
my_vector_sequence[my_vector_sequence <=12]
```
# Data Matrices

```{r}
Philosophers_Stone <- c(317.5, 657.1)
Chamber_of_Secrets <- c(261.9, 616.9)
Prisoner_of_Azkaban <- c(249.5, 547.1)
Goblet_of_Fire <- c(290.0, 606.8)
Order_of_the_Phoenix <- c(292.0, 647.8)
Half_Blood_Prince <- c(301.9, 632.4)
Deathly_Hallows_1 <- c(295.9, 664.3)
Deathly_Hallows_2 <- c(381.0, 960.5)

box_office <- c(Philosophers_Stone, Chamber_of_Secrets, Prisoner_of_Azkaban, Goblet_of_Fire, Order_of_the_Phoenix, Half_Blood_Prince, Deathly_Hallows_1, Deathly_Hallows_2)
box_office

```
```{r}
harry_potter_matrix <- matrix(box_office, nrow = 8, byrow = T)
harry_potter_matrix
```
```{r}
region <- c("US", "non-US")
region
titles <- c("Philosophers_Stone", "Chamber_of_Secrets", "Prisoner_of_Azkaban", "Goblet_of_Fire", "Order_of_the_Phoenix", "Half_Blood_Prince", "Deathly_Hallows_1", "Deathly_Hallows_2")
titles

```
```{r}
colnames(harry_potter_matrix) <- region
rownames(harry_potter_matrix) <- titles
harry_potter_matrix
```
```{r}
global <- rowSums(harry_potter_matrix)
global

```
```{r}
all_harry_potter_matrix <- cbind(harry_potter_matrix, global)
all_harry_potter_matrix
```

## Practice

##### 1. What are the total earnings for the US and non-US regions?

```{r}
Na <- NA
us_total <- all_harry_potter_matrix[,1]
US_Gross <- c(sum(us_total),Na,Na)
nus_total <- all_harry_potter_matrix[,2]
Non_US_Gross <- c(Na,sum(nus_total),Na)
Global_gross <- c(Na,Na, sum(global))
jk_rowling_lul_matrix <- rbind(all_harry_potter_matrix,US_Gross , Non_US_Gross, Global_gross)
jk_rowling_lul_matrix
```

# A Few Extras

```{r}
harry_potter_matrix[2,1]
harry_potter_matrix[1:4]
non_us_earnings <- all_harry_potter_matrix[ ,2]
mean(non_us_earnings)
```

