---
title: "Homework3_AH"
author: "Aryss Hearne"
date: "1/12/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

##### **1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R.**

```{r}
help(msleep)
glimpse(msleep)
```
##### **2. Store these data into a new data frame `sleep`.**

```{r}
sleep <- data.frame(msleep)
sleep
```
##### **3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the code that you used to determine this below.**
```{r}
dim(msleep)
```
*Looks like the frame is 83 rows by 11 columns. Looking above, this seems to check out.* 

##### **4. Are there any NAs in the data? How did you determine this? Please show your code.**

```{r}
anyNA(sleep)
```
*Yes, there are several NAs present in the matrix.*

```{r}
is.na(sleep)
```
*WOOF. A TON of them.*

##### **5. Show a list of the column names is this data frame.**
```{r}
names(sleep)
```

##### **6. How many herbivores are represented in the data?**

```{r}
herbivores <- grepl("herbi", sleep$vore)
sum(herbivores, na.rm=T)
```
*There are 32 herbivores listed*

##### **7. We are interested in two groups; small and large mammals. Let’s define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.**

```{r}
tiny <- subset(sleep, bodywt<=1.0)
beefy <- subset(sleep, bodywt>=200)
bittybabies <- data.frame(tiny)
bittybabies
```
```{r}
chonkybois <- data.frame(beefy)
chonkybois
```
##### **8. What is the mean weight for both the small and large mammals?**

```{r}
weighin_bittyboi <- (bittybabies$bodywt)
mean(weighin_bittyboi)
```
*Looks like the smaller animals weigh, on average. ~.3kgs*

```{r}
weighin_chonkyboi <- (chonkybois$bodywt)
mean(weighin_chonkyboi)
```
*And, on average, the larger animals weigh ~1750kgs.*

##### **9. Using a similar approach as above, do large or small animals sleep longer on average?**

```{r}
sleepsmalls <- (bittybabies$sleep_total)
mean(sleepsmalls)
bignaps <- (chonkybois$sleep_total)
mean(bignaps)
```
*Looks like the tiny guys need more sleep.* 

##### **10. Which animal is the sleepiest among the entire dataframe?**

```{r}
sleepiest_bbs <- subset(sleep, sleep_total>=19)
neednaps <- data.frame(sleepiest_bbs)
neednaps
```

*Looks like the sleepiest lil guy is the little brown bat!* 

