---
title: "R_warmup_6_1"
author: "Aryss Hearne"
date: "1/21/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
library("janitor")
```
```{r}
#install.packages("skimr")
library("skimr")
```


```{r}
?msleep
names(msleep)
```
```{r}
remotes::install_github("allisonhorst/palmerpenguins")
```
# dplyr Practice
#### 1. Let’s do a bit more practice to make sure that we understand `select()`, `filter()`, and `mutate()`. Start by building a new data frame `msleep24` from the `msleep` data that: contains the `name` and `vore` variables along with a new column called `sleep_total_24` which is the amount of time a species sleeps expressed as a proportion of a `24-hour day`. Remove any rows with NA’s and restrict the `sleep_total_24` values to less than 0.3. Arrange the output in descending order.

```{r}
msleep%>%
  mutate(sleep_total_24=sleep_total/24)%>%
  select(name, vore, sleep_total_24)%>%
  filter(sleep_total_24<=0.3)%>%
  na.omit(msleep)%>%
  arrange(desc(sleep_total_24))
```

```{r}
summary(msleep)
```
```{r}
skim(msleep)
```
#Practice
#### Which taxonomic orders have species that belong to more than one class of `vore`? 


```{r}
msleep%>%
  select(name, genus, order, vore)%>%
  filter(vore!="herbi"&vore!="carni")
```

*I dunno - all of these species only have one class of vore - I'm not sure what this question is asking*

```{r}
head(msleep)
```
```{r}
large<-
  msleep%>%
  select(name, genus, bodywt, sleep_total)%>%
  filter(bodywt>200)%>%
  arrange(desc(bodywt))
large
```
```{r}
mean(large$sleep_total)
```

```{r}
msleep %>% 
  filter(bodywt > 200) %>%
  summarize(mean_sleep_lg = mean(sleep_total))
```
```{r}
msleep %>% 
    filter(bodywt > 200) %>% 
    summarize(mean_sleep_lg = mean(sleep_total), 
              min_sleep_lg = min(sleep_total),
              max_sleep_lg = max(sleep_total),
              total = n())
```

```{r}
msleep%>%
  filter(order=="Primates")%>%
  summarize(mean_bodyweight = mean(bodywt),
            min_bodyweight = min(bodywt),
            max_bodyweight = max(bodywt),
            total=n())
```
```{r}
msleep%>%
  filter(bodywt>100)%>%
  summarise(n_genera=n_distinct(genus))
```
#Practice
#### 1. How many genera are represented in the msleep data frame?

```{r}
msleep%>%
  summarise(n_genera=n_distinct(genus))
```
#### 2. What are the min, max, and mean `sleep_total` for all of the mammals? Be sure to include the total n.

```{r}
msleep%>%
  select(name, sleep_total)%>%
  summarize(mean_sleep=mean(sleep_total),
            max_sleep=max(sleep_total),
            min_sleep=min(sleep_total),
            total=n())
```

```{r}
msleep%>%
  group_by(vore)%>%
  summarize(min_bodywt=min(bodywt),
            max_bodywt=max(bodywt),
            mean_bodywt=mean(bodywt),
            total=n())
```

# Practice
#### 1. Calculate mean brain weight by taxonoimc order in the msleep data. 

```{r}
msleep%>%
  group_by(order)%>%
  summarize(big_brain_time=mean(brainwt),
            total=n())
```

#### 2. What does `NA` mean? How are NAs being treated by the summarize function? 
*NAs are being treated as function-canceling data.*

#### 3. Try running the code again, but this time add `na.rm=TRUE`. What is the problem with Cetacea? Compare this to Carnivora.

```{r}
msleep%>%
  group_by(order)%>%
  summarize(big_brain_time=mean(brainwt),
            total=n())

```

