---
title: "Lab 6_2 AH"
author: "Aryss Hearne"
date: "1/25/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "R Notebook"
output: html_notebook
---

```{r}
library("tidyverse")
library("janitor")
library("skimr")
library("palmerpenguins")
```

```{r}
glimpse(penguins)
```

```{r}
penguins%>%
  group_by(island)%>%
  summarize(mean_body_mass_g=mean(body_mass_g),
            n=n())
```

```{r}
penguins%>%
  group_by(island)%>%
  summarize(number_NAs=sum(is.na(body_mass_g)))
```

```{r}
penguins%>%
  filter(!is.na(body_mass_g))%>%
  group_by(island)%>%
  summarize(mean_body_mass_g=mean(body_mass_g),
            sd_body_mass_g=sd(body_mass_g),
            n=n())
```

```{r}
penguins%>%
  group_by(island,species)%>%
  summarize(n=n(),.groups='keep')
```

```{r}
penguins%>%
  count(island,sort=T)
```

```{r}
penguins%>%
  group_by(island)%>%
  summarize(n=n())
```

```{r}
penguins%>%
  count(island,species,sort=T)
```

```{r}
penguins%>%
  tabyl(species,island)
```
```{r}
penguins%>%
  tabyl(species,island)%>%
  adorn_percentages()%>%
  adorn_pct_formatting(digits=1)%>%
  adorn_ns()
```

#Practice
#### Produce a summary of the mean for bill_length_mm, bill_depth_mm, flipper_length_mm, and body_mass_g within Adelie penguins only. Be sure to provide the number of samples.
```{r}
names(penguins)
```

```{r}
penguins%>%
  filter(species=="Adelie")%>%
  summarize(mean_bill_length_mm=mean(bill_length_mm, na.rm=T),
            sd_bill_length_mm=sd(bill_length_mm, na.rm=T),
            penguin_flippers=mean(flipper_length_mm, na.rm=T),
            n_samples=n())
```

#### How does the mean of `bill_length_mm` compare between penguin species? 

```{r}
penguins%>%
  group_by(species)%>%
  summarize(mean_bill_length_mm_2=mean(bill_length_mm, sort=T, na.rm=T),
            n_samples=n())
```

#### For some penguins, their sex is listed as `NA`. Where do these penguins occur?

```{r}
penguins%>%
  group_by(island)%>%
  summarize(number_NAs=sum(is.na(sex)),)
```

```{r}
penguins%>%
  summarize(distinct_species=n_distinct(species),
            distinct_island=n_distinct(island),
            distinct_sex=n_distinct(sex))
```
```{r}
penguins%>%
  summarize(across(c(species,island,sex),n_distinct))
```

```{r}
penguins%>%
  summarize(across(contains("mm"),mean,na.rm=T))
```

```{r}
penguins%>%
  group_by(sex)%>%
  summarize(across(contains("mm"), mean, na.rm=T))
```
```{r}
penguins%>%
  summarise_all(n_distinct)
```

```{r}
penguins%>%
  summarize(across(!c(species,island,sex),
                   n_distinct))
```
```{r}
penguins%>%
  summarise(across(starts_with("bill"),n_distinct))
```

#Practice
#### Produce separate summaries of the mean and standard deviation for bill_length_mm, bill_depth_mm, and fliper_length_mm for each penguin species. Be sure to provide the number of samples. 

```{r}
penguins%>%
  group_by(species)%>%
  summarize(across(contains("mm"), mean, na.rm=T),
            n=n())
```

```{r}
penguins%>%
  group_by(species)%>%
  summarize(across(contains("mm"), sd, na.rm=T),
            n=n())
```

